---
title: "FinalProjectWagnerKovscek"
author: "Grace Wagner and Nathanael Kovscek"
date: "2023-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
```{r, message= FALSE}
library(tidyverse)
library(ggplot2)
library(lubridate)

CODGameModes <- read.csv("~/Documents/github/STAT380/data/CODGameModes.csv")
CODGames_p1_380 <- read.csv("~/Documents/github/STAT380/data/CODGames_p1_380.csv")
CODGames_p2_380 <- read.csv("~/Documents/github/STAT380/data/CODGames_p2_380.csv")
CODMaps <- read.csv("~/Documents/github/STAT380/data/CODMaps.csv")
```

## Task 1 Data Cleaning and Data Visualization

Research Question: Which maps are the most likely to win the map vote?

```{r}
all(colnames(CODGames_p1_380) == colnames(CODGames_p2_380))
```


How I Plan to Answer this Question:

I will first make sure I append the CODGames_p1_380 and CODGames_p2_380 datasets, making sure to keep all of the observations, to make a complete dataset of the votes and maps selected. Then I will identify which votes ended in a tie. I will store this data separaetly from the data where a normal vote took place since Map1 is always chosen when there is a tie. 
To do this I plan to make another dataset that is made from the filtered out the rows that have no values in the MapVote column. Next I will identify how many times each map was chosen in a tie by summarizing the data and grouping by Choice making sure to add a N() value so that the number of times each Map was chosen is in the summary. I will repeat the summary and group_by steps for the dataset containing of regular votes. From there I will take the information from the two summaries, combine them, and make a visualization that shwos the results. To represent the results from ties, and from regular votes, I want to make sure that it is shown which maps were chosen from ties. To do this, I would like to make a barchart that is made of two colors and a key, one color will represnt regular votes and the other will represent results from tied votes.

```{r}
#join datasets
all_games <- CODGames_p1_380 %>% bind_rows(CODGames_p2_380)

# Check number of columns
# count(all_games) == (count(CODGames_p1_380) + count(CODGames_p2_380))

#filter out ties
map_stats <- all_games %>% 
  select(Map1, Map2, Choice, MapVote) %>% 
  filter(MapVote != "")
```

```{r}
map_stats$Choice[map_stats$Choice == "APocalypse"] <- "Apocalypse"
map_stats$Choice[map_stats$Choice == "Apocolypse"] <- "Apocalypse"
map_stats$Choice[map_stats$Choice == "Drive-In"] <- "Drive-in"
map_stats$Choice[map_stats$Choice == "Collateral"] <- "Collateral Strike"
map_stats$Choice[map_stats$Choice == "Collaterel Strike"] <- "Collateral Strike"
map_stats$Choice[map_stats$Choice == "Deisel"] <- "Diesel"
map_stats$Choice[map_stats$Choice == "Nuketown '84 Halloween"] <- "Nuketown '84"

CODMaps <- CODMaps %>% separate(Date, into = c("month", "day", "year"), sep = "/", convert = TRUE)
CODMaps <- CODMaps %>% 
  rename(Choice = Name)

CODMaps$year[CODMaps$year == "20"] <-  2020
CODMaps$year[CODMaps$year == "21"] <-  2021
CODMaps$year[CODMaps$year == "22"] <-  2022
  

# Join CODMaps so season variables are present for analysis
map_stats <- map_stats %>% inner_join(CODMaps, by = "Choice")
```

```{r}

ggplot(data = map_stats, mapping = aes(x = reorder(Choice, year), fill = as.factor(year))) + 
  geom_bar() + 
  labs(title = "Chosen Maps and their Selection Frequency",
       y = "Frequency",
       x = "Map Choice",
       fill = "Map Release Year") +
  theme(axis.text.x = element_text(angle = 90))

```
```{r}
x <- all_games %>% group_by(PrimaryWeapon) %>% 
  summarize(
    n = n()
  )
x$n <- x$n[order(x$n)]
x
```

