---
title: "Mini Project 3"
author: "Grace Wagner and Nathanael Kovscek"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
```{r warnings = FALSE, message = FALSE}
remove(list = ls())
# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(lubridate)

# Libraries for trees
library(rpart)
library(rattle)

#Part 1
library(readxl)
CODGames2_mp <- read_excel("data/CODGames2_mp.xlsx")
```
<<<<<<< HEAD
=======
<<<<<<< HEAD
>>>>>>> main

## Task 1

First, we would like to see the distribution of Full and Partial games. That way, we are able to see what the dimensions of the filtered data table should have.
<<<<<<< HEAD

=======
>>>>>>> main
```{r}
CODGames2_mp %>% 
  group_by(FullPartial) %>%
  summarise(n = n())
```

Given this output, when we filter out Partial values from the data table, we should have 191 observations.

```{r}
full_games <- CODGames2_mp %>%
  filter(FullPartial == "Full")
```

Now that this is done, let's see what the distribution of TotalXP is in Full Games. We will visualize this distribution using two boxplots where each is one of the two XPTypes that it is possible for the player to obtain.

```{r}
ggplot(full_games, mapping = aes(x = XPType, y = TotalXP)) +
  geom_boxplot()

```

This shown, it is much more favorable for the user to have the Double XP + 10% XPType than the 10% Boost type. The user's totalXP median earning is approximately 6000 more than it otherwise would be. Additionally, the interquartile range for Double XP + 10% is larger and there are more outliers on the high end than there is for 10% Boost type.

Out of curiosity, we would like to see the distribution of the XPTypes.

```{r}
CODGames2_mp %>% 
  group_by(XPType) %>%
  summarise(n = n())
```

I'm not sure if this is more or less surprising given what we know about the game, but my intuition suggests that something which increases the amount of earned XP by a median average of around 6000 would be more rare in the game. Additionally, even though there are fewer uses of the DoubleXP, we would still expect the larger IQR because the XP earned is double that which was received. Looking at the boxplots again, I would suggest that the IQR for the 10% Boost is half the IQR for the DoubleXP type!


## Task 2
<<<<<<< HEAD
total XP, eliminations, deaths, damage, XPType, and whether the player’s team won,

```{r}
tdm <- full_games %>% 
  filter(GameType == "HC - TDM") %>% 
  separate(Result, into = c("PlayersTeam", "OtherTeam"), sep = "-", convert = TRUE)

tdm <- tdm %>% 
  mutate(GameResult = tdm$PlayersTeam - tdm$OtherTeam) %>%
  mutate(GameResult2 = case_when(GameResult > 0 ~ "win", GameResult < 0 ~ "lose", GameResult == 0 ~ "tie")) %>%
  mutate(Win = ifelse(GameResult2 == "win", 1, 0)) %>% 
  select(TotalXP, Eliminations, Deaths, Damage, XPType, Win, Score)
```

### Part a

### Part b

Build a regression tree for predicting Score using total XP, eliminations, deaths, damage, XPType, and whether the player’s team won. Specify that each node must contain at least 15 observations. Display the tree and report the variables associated with the 3 highest variable importance values. (Include the variable importance values when mentioning the variables.)

```{r}
res <- rpart(Score ~ TotalXP + Eliminations + Deaths + Damage + XPType + Win, method="anova", data=tdm,
             minbucket=15)

#Create plot using fancyRpartPlot for rattle library
fancyRpartPlot(res, cex=0.8)
```
The variables with the three highest importance values are Damage (`r res[["variable.importance"]][["Damage"]]`), Eliminations (`r res[["variable.importance"]][["Eliminations"]]`), and TotalXP (`r res[["variable.importance"]][["TotalXP"]]`)

### Part c
=======

=======
>>>>>>> Grace
>>>>>>> main
