---
title: "Mini-Project"
author: "Grace Wagner and Nathanael Kovscek"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter Including Task 1
```{r warnings = FALSE, message = FALSE}
# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(lubridate)

#Part 1
library(readxl)
CODGames2_mp <- read_excel("data/CODGames2_mp.xlsx")
```

## Task 2

Explore 6 variables (at least 1 categorical and 1 quantitative). <br>

Identify:
- variable and type of variable #DONE
- types of values the variable can take #DONE
- amount of missing data within variable #DONE
- summary statistics #DONE
- visualization #DONE
- write a few sentences to describe what you have learned about the variable #NEED TO FINISH <br>

If Categorical:
- make a table of possible values and include counts or proportions #DONE
- make an appropriate plot to show the distribution of values #DONE <br>

if Quantitative:
- Include some basic summary stats like the mean, median, standard deviation, min, max, etc. #DONE
- make an appropriate plot to show the distribution of values #DONE


```{r}
#Variable 1: Eliminations
#type
typeof(CODGames2_mp$Eliminations)
#range
range(CODGames2_mp$Eliminations)
#missing value
sum(is.na(CODGames2_mp$Eliminations))
#summary stats
CODGames2_mp %>%
  summarize(
    mean = mean(Eliminations),
    median = median(Eliminations),
    sd = sd(Eliminations),
    min = min(Eliminations),
    max = max(Eliminations)
  )
#plot
hist(CODGames2_mp$Eliminations,
     xlab = "Eliminations",
     main = "Distribution of Eliminations")

#Variable 2: Deaths
typeof(CODGames2_mp$Deaths)
range(CODGames2_mp$Deaths)
sum(is.na(CODGames2_mp$Deaths))
CODGames2_mp %>%
  summarize(
    mean = mean(Deaths),
    median = median(Deaths),
    sd = sd(Deaths),
    min = min(Deaths),
    max = max(Deaths)
  )
hist(CODGames2_mp$Deaths,
     xlab = "Deaths",
     main = "Distribution of Deaths")

#Variable 3: Score
typeof(CODGames2_mp$Score)
range(CODGames2_mp$Score)
sum(is.na(CODGames2_mp$Score))
CODGames2_mp %>%
  summarize(
    mean = mean(Score),
    median = median(Score),
    sd = sd(Score),
    min = min(Score),
    max = max(Score)
  )
hist(CODGames2_mp$Score,
     xlab = "Score",
     main = "Distribution of Score")

#Variable 4: Damage
typeof(CODGames2_mp$Damage)
range(CODGames2_mp$Damage)
sum(is.na(CODGames2_mp$Damage))
CODGames2_mp %>%
  summarize(
    mean = mean(Damage),
    median = median(Damage),
    sd = sd(Damage),
    min = min(Damage),
    max = max(Damage)
  )
hist(CODGames2_mp$Damage,
     xlab = "Damage",
     main = "Distribution of Damage")

#Variable 5: TotalXP
typeof(CODGames2_mp$TotalXP)
range(CODGames2_mp$TotalXP)
sum(is.na(CODGames2_mp$TotalXP))
CODGames2_mp %>%
  summarize(
    mean = mean(TotalXP),
    median = median(TotalXP),
    sd = sd(TotalXP),
    min = min(TotalXP),
    max = max(TotalXP)
  )
hist(CODGames2_mp$TotalXP,
     xlab = "TotalXP",
     main = "Distribution of TotalXP")

#Number 6 Candidates (or remove a quantitative one and add both categorical)

#Variable 6: PrimaryWeapon
typeof(CODGames2_mp$PrimaryWeapon)
unique(CODGames2_mp$PrimaryWeapon)
sum(is.na(CODGames2_mp$PrimaryWeapon))
tableWeapons <- CODGames2_mp %>%
  group_by(PrimaryWeapon) %>%
  summarize(N = n())

ggplot(CODGames2_mp, aes(PrimaryWeapon)) +
  geom_bar() +
  labs(title = "Distribution of Primary Weapons", x = "Primaray Weapons")
  

#Variable 6: Choice #Has NA values
typeof(CODGames2_mp$Choice)
unique(CODGames2_mp$Choice)
sum(is.na(CODGames2_mp$Choice))
CODGames2_mp %>%
  group_by(Choice) %>%
  summarize(N = n())

ggplot(CODGames2_mp, aes(Choice)) +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 4)) +
  labs(title = "Distribution of Choices", xlab = "Choices")
```

## Task 3

All of our questions are regarding what the variables mean. Here they are: <br><br>
<ul>
  <li>What aspects of the game play into the variable Score? How can one score high? How does it differ from                  TotalXP?</li>
  <li>What would cause the variable "Choice" - which is the choice of map in which the game is played - to have an NA         value?</li>
  <li>What is the relationship between Map1, Map2, Choice, and MapVote?</li>
  <li>What does the ratio mean in the variable MapVote?</li>
  <li>What are the different GameTypes?</li>
  <li>What does the Results table represent and does the meaning change among game types?</li>
  <li>What game mode is HC - TDM? Specifically, what does TDM stand for?</li>
</ul>

## Task 4

### Part 1

Is the player’s performance, as quantified by the amount of experience points gained (TotalXP variable) changing over time?

```{r}
# Create a new table with TotalXP, Date, and then year, month, day derived from Date
# the lubridate package is loaded in with tidyverse

xp <- CODGames2_mp %>% 
  select("Date", "TotalXP") %>% 
  mutate(year = year(CODGames2_mp$Date), 
         month = month.name[month(CODGames2_mp$Date)],
         day = day(CODGames2_mp$Date)
         )

# Use fct_inorder to order boxplots by month
ggplot(data = xp, mapping = aes( x = fct_inorder(month), y = TotalXP)) +
  geom_boxplot() +
  labs(title = "TotalXP Earning by Month", 
       x = "Month",
       y = "TotalXP Earned")

```
<br> *Assessment:* <br><br>

It does not appear that there is any improvement whatsoever across the three months in which the user played the game. In fact, the median value of XP earned in a day during a given month is less in both July and August as compared to June. Arguably, the user's performance worsened.<br><br>

However, there are numerous variables at play here. Game play could have affected the TotalXP value. Additionally, the user might decide that, once his performance with a particular weapon reaches a certain level, he wants to use a different weapon. So, while he apparently is not earning more XP over time, that does not necessarily mean that he is becoming worse at the game.<br><br>

The one final note is that the user might not have played as much from month to month. Such variability in game play would also affect our assessment.<br><br>

### Part 2

question I was thinking we could ask: which primary weapon (or choice, but choice has NA values) resulted in the most deaths/elimintaions/totalXp


I am interested in answering the question: what is the relationship between deaths and eliminations?

```{r}
elims <- CODGames2_mp %>%
  select("Date", "Deaths", "Eliminations", "GameType") %>% 
    mutate(year = year(CODGames2_mp$Date), 
         month = month.name[month(CODGames2_mp$Date)],
         day = day(CODGames2_mp$Date)
         )
  
ggplot(data = elims, mapping = aes(x = Eliminations, y = Deaths, color = GameType)) +
  geom_point() + 
  labs(title = "Elimations Compared to Deaths by Game Mode",
       y = "Times User Died",
       x = "Times User Eliminated an Opponent"
       )
```
After observing this visualization, a couple things become apparent. One is that the game mode has a major influence on the number of deaths and eliminations that the user can get in a game. It would appear that TDM is the game mode that is most preferred by our user, but it is also the game which seems to have the lowest number of deaths and elimations. The Hardpoint mode is what appears to accumulate the most deaths and eliminations. <br><br>

Because my main question asks about the relationship between deaths and eliminations by game, and that three out of the four game modes have few representations, I will remove the three game modes that aren't TDM and then check out the relationship between our two desired variables.

```{r}
# Table filtering to only have the necessary values
elims_TDM <- CODGames2_mp %>%
  select("Date", "Deaths", "Eliminations", "GameType", "PrimaryWeapon") %>% 
    mutate(year = year(CODGames2_mp$Date), 
         month = month.name[month(CODGames2_mp$Date)],
         day = day(CODGames2_mp$Date)
         ) %>% 
  filter(GameType == "HC - TDM")

## Visualization
ggplot(data = elims_TDM, mapping = aes(x = Eliminations, y = Deaths)) +
  geom_point() + geom_smooth(method = lm)
  labs(title = "Elimations Compared to Deaths by Game Mode",
       y = "Times User Died",
       x = "Times User Eliminated an Opponent"
       )
```
<br>
So, as it turns out, there is actually no relationship between the number of deaths and eliminations in the game mode TDM - the most commonly played game mode by our user. This is very surprising because, typically, games such as the one which we have data for will have a little bit of time you have to wait in which to respawn. The hypothesis at the beginning was that more deaths means more time waiting to respawn which means less time to accrue eliminations. However, the data shows that this does not appear to be the case! <br>

There is, of course, another explanation. That is, our assumptions about how the game works is incorrect. There might not be any time between being killed and being allowed to respawn into the game. This, then, would be a good question to ask and is why it ended up being included in the section regarding our questions about the data features. <br>

### Extra: Part 3

There is one final thing which we would like to explore. We are curious to see if any relationships appear if we add in an additional layer of complexity to the above analysis. Namely, does the choice of primary weapon influence the number of deaths and eliminations in a TDM game. Because there are eight different choices of primary weapon, we only looked into the top three most used. Those would be: M-16, MP-5, and the Pelington 703. This is explored below.

```{r}
elims_TDM_topguns <- CODGames2_mp %>%
  select("Date", "Deaths", "Eliminations", "GameType", "PrimaryWeapon") %>% 
    mutate(year = year(CODGames2_mp$Date), 
         month = month.name[month(CODGames2_mp$Date)],
         day = day(CODGames2_mp$Date)
         ) %>% 
  filter(GameType == "HC - TDM",
         PrimaryWeapon == "M16"| PrimaryWeapon == "MP5"|PrimaryWeapon == "Pelington 703")

## Visualization
ggplot(data = elims_TDM_topguns, mapping = aes(x = Eliminations, y = Deaths,)) +
  geom_point() +
  geom_smooth(method = lm,
              color = "blue") +
  labs(title = "Elimations Compared to Deaths by Gun",
       subtitle = "TDM Game Mode",
       y = "Times User Died",
       x = "Times User Eliminated an Opponent"
       ) +
  facet_grid(. ~ PrimaryWeapon) +
  theme_bw()
```
<br>
After this, while it doesn't seem that there is any relationship when the user is using the M16, there is an apparent negative relationship between Deaths and Eliminations when the user is using the MP5 and a positive relationship when the user is using the Pelington 703. <br>

Presently, we have no hypothesis as to why this might be. It is, however, an interesting thing and is fundamently why we explored it. :)

## Discussion of Group Participation

In this project, Grace and I did an even split of the work. She did questions 1 and 2 and I did questions 3 and 4. To do this effectively, we met at several points in the process to discuss not only this distribution of work, but also what visualizations to make for part 2 and what questions to ask for part 3. We also discussed what would be the most interesting question to explore for part four. We both found this distribution of work to be effective and would do something similar in the future. <br>
