---
title: "Mini-Project"
author: "Grace Wagner and Nathanael Kovscek"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter Including Part 1
```{r}
#Part 1
library(readxl)
CODGames2_mp <- read_excel("C:/Users/Grace Wagner/Downloads/CODGames2_mp.xlsx")
#This wont work for you will it since its a path for my computer? I did this for now so that I could work on part 2, but I know you mentioned using the data in github and I thought I knew what to do but once I looked at it further I realized I wasn't sure. 
#I also just merged the FrontMatter and part 1 because I figured since part 1 was importing data it would be nice to keep it in the Front matter.

library(dplyr)
library(ggplot2)
```

## Part 2
Explore 6 variables (at least 1 categorical and 1 quantitative). 
Identify:
- variable and type of variable #DONE
- types of values the variable can take #DONE
- amount of missing data within variable #DONE
- summary statistics #DONE
- visualization #DONE
- write a few sentences to describe what you have learned about the variable #NEED TO FINISH
If Categorical:
- make a table of possible values and include counts or proportions #DONE
- make an appropriate plot to show the distribution of values #DONE
if Quantitative:
- Include some basic summary stats like the mean, median, standard deviation, min, max, etc. #DONE
- make an appropriate plot to show the distribution of values #DONE
```{r}
#Variable 1: Eliminations
#type
typeof(CODGames2_mp$Eliminations)
#range
range(CODGames2_mp$Eliminations)
#missing value
sum(is.na(CODGames2_mp$Eliminations))
#summary stats
CODGames2_mp %>%
  summarize(
    mean = mean(Eliminations),
    median = median(Eliminations),
    sd = sd(Eliminations),
    min = min(Eliminations),
    max = max(Eliminations)
  )
#plot
hist(CODGames2_mp$Eliminations,
     xlab = "Eliminations",
     main = "Distribution of Eliminations")

#Variable 2: Deaths
typeof(CODGames2_mp$Deaths)
range(CODGames2_mp$Deaths)
sum(is.na(CODGames2_mp$Deaths))
CODGames2_mp %>%
  summarize(
    mean = mean(Deaths),
    median = median(Deaths),
    sd = sd(Deaths),
    min = min(Deaths),
    max = max(Deaths)
  )
hist(CODGames2_mp$Deaths,
     xlab = "Deaths",
     main = "Distribution of Deaths")

#Variable 3: Score
typeof(CODGames2_mp$Score)
range(CODGames2_mp$Score)
sum(is.na(CODGames2_mp$Score))
CODGames2_mp %>%
  summarize(
    mean = mean(Score),
    median = median(Score),
    sd = sd(Score),
    min = min(Score),
    max = max(Score)
  )
hist(CODGames2_mp$Score,
     xlab = "Score",
     main = "Distribution of Score")

#Variable 4: Damage
typeof(CODGames2_mp$Damage)
range(CODGames2_mp$Damage)
sum(is.na(CODGames2_mp$Damage))
CODGames2_mp %>%
  summarize(
    mean = mean(Damage),
    median = median(Damage),
    sd = sd(Damage),
    min = min(Damage),
    max = max(Damage)
  )
hist(CODGames2_mp$Damage,
     xlab = "Damage",
     main = "Distribution of Damage")

#Variable 5: TotalXP
typeof(CODGames2_mp$TotalXP)
range(CODGames2_mp$TotalXP)
sum(is.na(CODGames2_mp$TotalXP))
CODGames2_mp %>%
  summarize(
    mean = mean(TotalXP),
    median = median(TotalXP),
    sd = sd(TotalXP),
    min = min(TotalXP),
    max = max(TotalXP)
  )
hist(CODGames2_mp$TotalXP,
     xlab = "TotalXP",
     main = "Distribution of TotalXP")

#Number 6 Candidates (or remove a quantitative one and add both categorical)

#Variable 6: PrimaryWeapon
typeof(CODGames2_mp$PrimaryWeapon)
unique(CODGames2_mp$PrimaryWeapon)
sum(is.na(CODGames2_mp$PrimaryWeapon))
tableWeapons <- CODGames2_mp %>%
  group_by(PrimaryWeapon) %>%
  summarize(N = n())

ggplot(CODGames2_mp, aes(PrimaryWeapon)) +
  geom_bar() +
  labs(title = "Distribution of Primary Weapons", x = "Primaray Weapons")
  

#Variable 6: Choice #Has NA values
typeof(CODGames2_mp$Choice)
unique(CODGames2_mp$Choice)
sum(is.na(CODGames2_mp$Choice))
CODGames2_mp %>%
  group_by(Choice) %>%
  summarize(N = n())

ggplot(CODGames2_mp, aes(Choice)) +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(n.dodge = 4)) +
  labs(title = "Distribution of Choices", xlab = "Choices")
```

## Part 3
question I was thinking we could ask: which primary weapon (or choice, but choice has NA values) resulted in the most deaths/elimintaions/totalXp

## Part 4